The Impact of Disk Winds on the Optical Spectra of Cataclysmic Variables

### 2014-11-06 draft.tex
Some changes and added fig1_nl_spectra generation folder

Cataclysmic variables (CVs) are systems in which a white dwarf accretes matter from a donor
star via Roche-lobe overflow. In non-magnetic systems this accretion
is mediated by an accretion disk which forms around the white dwarf. 
Nova-like variables (NLs) are a particular subclass of objects in which the accretion disk
is in a permanent state of relatively high accretion rate 
($\dot{M} \sim 10^{-8}$~M$_{\odot}$~yr$^{-1}$).  This makes nova-like variables the obviousl
laboratory for studying the properties of steady-state accretion disks.
%making them the perfect laboratories to test the widely-used`$\alpha$-disk' model proposed by \cite{shakurasunyaev1973}.

For over three decades, it has been known that winds emanating from the accretion disk
are important in shaping the ultraviolet (UV) spectra of high-state CVs (Heap 1978), 
the most spectacular evidence being the P-Cygni like profiles of resonance lines such as 
\civfull\ (see e.g. Cordova \& Mason 1982\nocite{cordova1982}).  As discussed below, it is now possible to produce synthetic spectra 
of CVs with winds that are qualitatively very similar to those observed.  However, less is known about the 
the effects of the wind on the optical spectra of high state CVs, which are characterized by a relatively featureless 
continuum and  emission from the
Balmer lines of H, which in many cases are single peaked, especially, the SW Sex stars \citep{HSK86, DR95}, unlike the double-peaked lines one might expect from a disk.   Some work has been carried out.
% Despite this, in many types of systems the impact of winds on observational features is essentially unknown, and many theoretical questions relating to their origin  and driving mechanism remain unanswered.
Murray \& Chiang (1996) have argued that the single peaked emission lines arise due to radiative transfer effects in the wnd
% While the effect of the wind on the UV is at least partially clear, it may also have more subtle  effects across the spectral range. It has been proposed that an accretion disk wind can provide a mechanism for single-peaked emission lines (Murray \& Chiang 1996),  such as those seen in Nova-like variables and, especially, the SW Sex stars \citep{HSK86, DR95}.
It is particularly telling that single-peaked 
lines in high-inclination systems can be seen even in eclipse, suggesting 
that a spatially extensive wind may be the source of emission (see figure~\ref{novalikes}).
%%(Honeycutt, Schlegel, \& Kaitchuck 1986; Dhillon \& Rutten 1995). 
Further evidence of the influence of winds in the optical can occasionally 
be occasionally seen in P-Cygni like profiles observed in lines such as \ha\ 
and He \textsc{i} $\lambda5876$ \citep{RN98}.

@@

The continuum spectra of high disk-dominated CVs is generally assumed to be
dominated by emission from the diskand and is often modeled in terms of a set of stellar atmospheres
and/or disk models ({bf reference some of Linnells papers here}), using Hubeny's TLUSTY code {\cite{})
The optical continua observed from  high-state, disk-dominated CVs lack pronounced ``Balmer jumps'', 
spectra simulated from summed atmospheres.   One possible explanation for this problem is that disk atmospheres calculated
with TLUSTY simply do not represent the physics of disks properly ((see e.g. Shaviv et al. 1991; 
but also more recent work by Idan et al. 2010).
\nocite{idanshaviv2010} \nocite{shaviv1991}.   An altenative suggested by \cite{KLWB98} is that recombination continuum emission from a disk wind or `transition region'
fills  in the Balmer absorption edge.

@@

Here, we present work using the same Monte Carlo radiative transfer code as LK02, 
but incorporating additional `macro-atom' techniques suggested by Lucy (2002, 2003). 
Our aim is to model high-state CVs over a wide spectral range.
In particular, we hope to assess the potential contribution
to the optical spectrum from a disk wind model which preserves the UV features.
It will then be possible to see how closely
our models resemble the spectra in figure 1 as we change wind parameters.
We will investigate whether a wind can produce significant
recombination emission lines and whether the recombination
continuum emission is sufficient to mask the Balmer jump from a disk atmosphere.
We will also learn if the mechanism for single-peaked line emission
identified by \cite{MC96} operates  in  parameterized wind models capable of reproducing 
other spectral features of observed CVs.  The remainder of this report is organized as followd
In section 2 we present the code and provide a brief description of macro-atoms 
and the radiative transfer methods we employ.  In section 3, 
We describe the kinematics and geometry of our biconical wind model.
In section 4, we present spectra simulated from the benchmark model employed by LK02, and in section 5, we carry out 
aparameter search to determine how spectra change as a function of several of the parmeters in the model. In section 6, we present a
model which better reproduces the optical spectrum of a NL, and, 
finally, in section 7, we summarise our findings.

@@

\py is a Monte Carlo (MC) radiative transfer code which uses
the Sobolev approximation to treat line transfer. It operates 
in two distinct cycles. First, the ionization state, level populations
and temperature structure are calculated. This is done by
flying MC energy quanta (`photons') through a pre-defined wind geometry and recording
estimators that characterize the radiation field in each cell of the grid, before choosing  a new electron temperature 
that more closely balances heating and cooling in the plasma. Using the new electron temperature and the radiative estimators, we then
revise the ionization state of the wind.  
This process is repeated  a converged
solution is reached with heating and cooling balanced throughout the wind. Second, the spectrum over the desired spectral range is synthesized by tracking
the energy packets through the wind and computing the 
emergent spectra at a number of user-specified viewing angles.

\py is designed to operate in a number of different
regimes, both in terms of the scale of the system, and the 
shape and characteristic energies of the illuminating spectrum. 
It was originally developed by LK02 in order to model the UV spectra
of CVs with a simple biconical disk wind model.. 
Higginbottom et al.\nocite{higginbottom2013} (2013; hereafter H13) used the same code to
present a benchmark model for broad absorption line (BAL) QSOs.
As part of this effort, the ionization scheme was improved so that 
the code is now capable of treating arbritrary SEDs.
%the power-law X-ray spectrum exhibited by AGN.
% This involved using a spectral model the mean intensity in a cell
% computed from MC estimators for, instead of using a dilute blackbody
% approximation. Compton processes
% were also incorporated into the heating and cooling balance.
In addition, Sim et al. (2005; hereafter SDL05)\nocite{simmacro2005} used the code to model
Brackett and Pfund line profiles of H in young-stellar objects (YSOs) using \textsc{Python,} and as part of this effort 
implemented a `macro-atom' mode  in order to correctly treat recombination lines of H.

@@

Whereas SDL05 used \py itself to apply the macro-atom 
technique to a Hydrogen-only model, we
adopt a hybrid ionization and excitation scheme to model the spectra of CVs. 
Atomic pecies can be treated as either full macro-atoms
or using the two-level atom approximation.  This is  important in our multi-dimensional code to balance the need to calculate certain transitions in detail and the desire to minimize the set of
computational resources required.  For CVs, this allows us 
%In the latter case, they still follow the indivisible packet constraint.This allows one 
to preserve the fast treatment of the UV resonance lines while providing 
an improved representation of processes with more complex
transition paths such as recombination, H and He. %This is important when modeling features  such as the Balmer lines and also recombination {\sl continuum} emission.

@@

\py allows for several types of biconical wind models and various options for calculating the spectra of the components of the system.
Here we describe some of the features of the  biconical wind models used in this study of CVs.

@@

Photons originate from the accretion disk and a central WD in our model. It is also possible to include
a boundary layer with user-defined temperature and emitting area. Emission from the wind is dealt with
by reprocessing of the indivisible energy packets originally produced in the disk or central WD. {\bf ksl -Isn't the wind also allow to emit?  If so, I think saying the wind is dealt with by reporcessing is misleading.}

@@

\py has has some flexibility when treating the accretion disk as a source of photons. 
The disk is broken down into annuli 
such that each annulus contributes an equal amount to the bolometric luminosity. 
We adopt the temperature profile of a standard \cite{shakurasunyaev1973} $\alpha$-disk.
%shown in figure 1.
An annulus can then
be treated either as a blackbody with the corresponding effective temperature or as a stellar atmosphere model
with the appropriate surface gravity and effective temperature. 
Both of these methods have limitations. It is known
from studies of CV accretion disks that they exhibit absorption cores 
formed in an optically thick disk atmosphere.
This can impact the overall spectrum even if a wind is present 
\citep[see e.g.][]{dhillon1996}.
However, the stellar atmosphere models
have had difficulties reproducing the required spectral shape \citep{wade1988}, 
and there are inconsistencies when using a 
stellar atmosphere beneath a dense
wind region that itself goes some way to simulating the disk atmosphere. 
In addition, \cite{linnell2010} found that deviations from the predicted 
$T\propto r^{-3/4}$ temperature profile are required to fit the data
in certain cases.  {\bf ksl - I thought Wade said that both black bodies and stellar atmospheres had trouble reproducing the color of disks.  Please check}

@@

\begin{table}
\centering
\begin{tabular}{p{3cm}p{4cm}}
Model A \\
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{WD}$ 	 &	 $0.8 M_{\odot}$ \\ 
$R_{WD}$ 	 &	 $7\times10^{8}$cm\\ 
$\dot{M}_{acc}$ 	 &	 $10^{-8}~M_{\odot}yr^{-1}$\\ 
$\dot{M}_{wind}$  &	$10^{-9}~M_{\odot}yr^{-1}$å\\ 
$r_{min}$ 	&	 $4 R_{WD}$\\ 
$r_{max}$ 	&	 $12 R_{WD}$ \\ 
$\theta_{min}$ 	&	 $20.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $65.0^{\circ}$ \\ 
$\gamma$ 	&	 $1$ \\ 
$v_{\infty}$ 	&	 $3v_{esc}$ \\ 
%%$R_v$ 	        &	 $7\times10^{10}$cm \\ 
$R_v$ 	        &	 $100 R_{WD}$ \\ 
$\alpha$ 	&	 $1.5$ \\
\end{tabular}
\centering
\caption{
Parameters used for the geometry and kinematics of the benchmark CV model.}
\label{wind_param}
\end{table}

@@

\begin{figure} %fullpage
\includegraphics[width=\textwidth]{figures/fig5_uv_opt.eps}
\caption{
UV (left) and optical (right) synthetic spectra for model A, our benchmark model,
computed at sightlines of 10, 27.5, 45, 62.5 and 80 degrees.	
The inset plots show zoomed-in line profiles for 
\heiiuv\ and \ha. Double-peaked line emission can be seen in 
\heiiuv, \heiiopt, \ha\ and some He I lines, but the 
line emission is not always sufficient to overcome the absorption
cores from the stellar atmosphere models. The model
also produces a prominent \heiioptnew\ line at high inclinations,
}
\label{spec}
\end{figure} %fullpage\

Figure~\ref{spec} shows optical synthetic spectra for the benchmark 
model for five inclination angles ranging from $10^\circ$ to $80^\circ$ .
%5 different sightlines.  
The spectra have some lines of H and He which have emission components that are strong enough to overcome
the absorption cores in the underlying disk spectra. These emission lines are generally double peaked, 
as they are formed near the base of the wind where rotational velocity
can still dominate over the slowly increasing poloidal velocity. The line 
opacity is not yet high enough for the poloidal velocity gradient to 
cause single-peaked profiles.  {\bf ksl - So this last statement is based on information that has not yet been submitted as evidence. Be careful of the flow of the paper here.}
We also see a general trend from absorption to emission in spectral lines
with increasing inclination, as one would expect from our wind geometry.
This trend is observed in CVs and can be seen clearly in figure 1. 
% Note, however,
% that in our models the \heiiuv\ line does not exhibit this behaviour, which
% is somewhat surprising. {\bf NOTE: we need to understand this!}.

@@

The benchmark model presented above shows that a wind with the properties
of Model A, a model that successfully reproduces the observed UV spectral features
of CVs,  
would also leave a clear imprint on the optical spectrum. {\bf ksl - why are we introducing the terminology of Model A at this point}. It is important
to understand how changing the kinematics of the wind impacts on this
effect.   
A full parameter search is beyond the scope of this study, 
but we can examine the evolution
of the spectrum with a few key parameters. {\bf - ksl - I don't actually like the previous sentence.  We should some how say we want to study a few parameters that we think with help,
without saying that something is beyond the scope}  In particular, this
will help shed light on the predictions of, e.g., Murray \& Chiang (1996)
and Knigge et al. (1998) regarding the single-peaked H$\alpha$ emission
and absence of a pronounced Balmer absorption edge respectively.

@@

A full kinematic parameter search may prove instructive, but
is beyond the scope of this particular study. {\bf - ksl - you said this once already, and I don't think it should be explicitly stated at all}
We have therefore conducted a limited parameter 
search, where we focus on 
two variables in order to explore the trends in our model. 
These are:

@@

\begin{table}
\centering
\begin{tabular}{p{3cm}p{4cm}}
Model B \\
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{WD}$ 	 &	 $0.8 M_{\odot}$ \\ 
$\dot{M}_{acc}$ 	 &	 $10^{-9}~M_{\odot}yr^{-1}$\\ 
$\dot{M}_{wind}$  &	$10^{-8}~M_{\odot}yr^{-1}$\\ 
$r_{min}$ 	&	 $4 R_{WD}$\\ 
$r_{max}$ 	&	 $12 R_{WD}$ \\ 
$\theta_{min}$ 	&	 $20.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $65.0^{\circ}$ \\ 
$\gamma$ 	&	 $1$ \\ 
$v_{\infty}$ 	&	 $3v_{esc}$ \\ 
$R_v$ 	        &	 $142.9 R_{WD}$ \\ 
$\alpha$ 	&	 $4$ \\
\end{tabular}
\centering
\caption{Wind geometry parameters used in the next-generation CV model.}
\label{modelb}
\end{table}

With modest changes to the kinematics of the benchmark CV model,
it is possible to produce a model
that comes closer to reproducing the observed optical spectrum
of a NLs. The parameters for this model (imaginatively dubbed Model B)
are show in table~\ref{modelb}, and the synthetic spectrum
is shown in figure~\ref{uvoptb}. An in and out of eclipse comparison 
between the synthetic spectrum
at an angle of $80^\circ$ and the high-inclination NL RW Tri 
is also shown in figure~\ref{rwtricomp}. The inclination
for RW Tri has been disputed, with claimed values of $75^\circ$ \citep{groot2004},
$70.5^\circ$ \citep{smak1995}, $80^\circ$ \citep{longmore1981} and 
$82^\circ$\citep{frankking1981}. A lower inclination
would mean that our model would simply require lower densities 
and smaller emission regions to produce the required continuum level
and single peaked emission, so our conclusions from this comparison are
not particular sensitive to the value chosen.
We also emphasize that this is no sense a fit to the data. The model
is purely designed to reproduce the general behaviour of NLs
and give an indication of the likely densities and kinematics 
required in the wind.
%quotes a figure of , whereas  halevinhenden2008

The similarity with the observed spectrum is striking. 
Model B produces strong emission in all the Balmer lines, 
with line to continuum ratios
comparable to those in RW Tri. 
The line to continuum ratio increases during eclipse,
as one expects if the emission is produced in a disk wind, 
thereby mimicking the behaviour of SW Sex stars.
The single-peaked \ha\ line is produced by our models with a 
very similar width to the observed spectrum, although it is perhaps slightly more
`flat-topped' in the synthesized data.

@@

\begin{figure} %fullpage
\includegraphics[width=0.8\textwidth]{figures/fig13_eclipse.eps}
\caption{{\sl Top Panel:} In and out of eclipse spectra of the high
inclination NL RW Tri. {\sl Bottom Panel:} In and out of eclipse synthetic
spectra from Model B.
The artificial `absorption' feature just redward of the Balmer jump
is caused for the reasons described in section 5.2.}
\label{rwtricomp}
\end{figure} %fullpage

>>>>

Cataclysmic variables (CVs) are systems in which a white dwarf accretes matter from a donor
star via Roche-lobe overflow. In non-magnetic systems this accretion
is mediated by an accretion disk which forms around the white dwarf. 
Nova-like variables (NLs) are a particular subclass of objects in which the accretion disk
is in a permanent state of relatively high accretion rate 
($\dot{M} \sim 10^{-8}$~M$_{\odot}$~yr$^{-1}$).  This makes nova-like variables the obvious
laboratory for studying the properties of steady-state accretion disks.
%making them the perfect laboratories to test the widely-used`$\alpha$-disk' model proposed by \cite{shakurasunyaev1973}.

For over three decades, it has been known that winds emanating from the accretion disk
are important in shaping the ultraviolet (UV) spectra of high-state CVs (Heap 1978), 
the most spectacular evidence being the P-Cygni like profiles of resonance lines such as 
\civfull\ (see e.g. Cordova \& Mason 1982\nocite{cordova1982}).  
As discussed below, it is now possible to produce synthetic spectra 
of CVs with winds that are qualitatively very similar to those observed.  
However, less is known about the the effects of the wind on the optical spectra of high state CVs, 
which are characterized by a relatively featureless continuum and  emission from the
Balmer lines of H, which in many cases are single peaked, especially, the SW Sex stars \citep{HSK86, DR95}, unlike the double-peaked lines one might expect from a disk.   Some work has been carried out.
% Despite this, in many types of systems the impact of winds on observational features is essentially unknown, and many theoretical questions relating to their origin  and driving mechanism remain unanswered.
Murray \& Chiang (1996) have argued that the single peaked emission lines arise 
due to radiative transfer effects in the wind.
% While the effect of the wind on the UV is at least partially clear, it may also have more subtle  effects across the spectral range. It has been proposed that an accretion disk wind can provide a mechanism for single-peaked emission lines (Murray \& Chiang 1996),  such as those seen in Nova-like variables and, especially, the SW Sex stars \citep{HSK86, DR95}.
It is particularly telling that single-peaked 
lines in high-inclination systems can be seen even in eclipse, suggesting 
that a spatially extensive wind may be the source of emission (see figure~\ref{novalikes}).
%%(Honeycutt, Schlegel, \& Kaitchuck 1986; Dhillon \& Rutten 1995). 
Further evidence of the influence of winds in the optical can occasionally 
be occasionally seen in P-Cygni like profiles observed in lines such as \ha\ 
and He \textsc{i} $\lambda5876$ \citep{RN98}.

@@

The continuum spectra of high disk-dominated CVs is generally assumed to be
dominated by emission from the disk and and is often modeled in terms of a set of stellar atmospheres
and/or disk models ({bf reference some of Linnells papers here}), using Hubeny's TLUSTY code {\cite{})
The optical continua observed from  high-state, disk-dominated CVs lack pronounced ``Balmer jumps'', 
spectra simulated from summed atmospheres.  One possible explanation for this problem is that disk atmospheres calculated
with TLUSTY simply do not represent the physics of disks properly 
((see e.g. Shaviv et al. 1991; but also more recent work by Idan et al. 2010).
\nocite{idanshaviv2010} \nocite{shaviv1991}.   
An alternative suggested by \cite{KLWB98} is that recombination continuum emission 
from a disk wind or `transition region' fills  in the Balmer absorption edge.

@@

Here, we present work using the same Monte Carlo radiative transfer code as LK02, 
but incorporating additional `macro-atom' techniques suggested by Lucy (2002, 2003). 
Our aim is to model high-state CVs over a wide spectral range.
In particular, we hope to assess the potential contribution
to the optical spectrum from a disk wind model which preserves the UV features.
It will then be possible to see how closely
our models resemble the spectra in figure 1 as we change wind parameters.
We will investigate whether a wind can produce significant
recombination emission lines and whether the recombination
continuum emission is sufficient to mask the Balmer jump from a disk atmosphere.
We will also learn if the mechanism for single-peaked line emission
identified by \cite{MC96} operates  in  parameterized wind models capable of reproducing 
other spectral features of observed CVs.  The remainder of this report is organized as follows.
In section 2 we present the code and provide a brief description of macro-atoms 
and the radiative transfer methods we employ.  In section 3, 
we describe the kinematics and geometry of our biconical wind model.
In section 4, we present spectra simulated from the benchmark model employed by LK02, and in section 5, we carry out 
aparameter search to determine how spectra change as a function of several of the parmeters in the model. In section 6, we present a
model which better reproduces the optical spectrum of a NL, and, 
finally, in section 7, we summarise our findings.

@@

\py is a Monte Carlo (MC) radiative transfer code which uses
the Sobolev approximation to treat line transfer. It operates 
in two distinct cycles. First, the ionization state, level populations
and temperature structure are calculated. This is done by
flying MC energy quanta (`photons') through a pre-defined wind geometry and recording
estimators that characterize the radiation field in each cell of the grid, before choosing  a new electron temperature 
that more closely balances heating and cooling in the plasma. Using the new electron temperature and the radiative estimators, we then
revise the ionization state of the wind.  
This process is repeated until a converged
solution is reached with heating and cooling balanced throughout the wind. 
Second, the spectrum over the desired spectral range is synthesized by tracking
the energy packets through the wind and computing the 
emergent spectra at a number of user-specified viewing angles.

\py is designed to operate in a number of different
regimes, both in terms of the scale of the system, and the 
shape and characteristic energies of the illuminating spectrum. 
It was originally developed by LK02 in order to model the UV spectra
of CVs with a simple biconical disk wind model.
Higginbottom et al.\nocite{higginbottom2013} (2013; hereafter H13) used the same code to
present a benchmark model for broad absorption line (BAL) QSOs.
As part of this effort, the ionization scheme was improved so that 
the code is now capable of treating arbritrary SEDs.
%the power-law X-ray spectrum exhibited by AGN.
% This involved using a spectral model the mean intensity in a cell
% computed from MC estimators for, instead of using a dilute blackbody
% approximation. Compton processes
% were also incorporated into the heating and cooling balance.
In addition, Sim et al. (2005; hereafter SDL05)\nocite{simmacro2005} used the code to model
Brackett and Pfund line profiles of H in young-stellar objects (YSOs) using \textsc{Python,} and as part of this effort 
implemented a `macro-atom' mode  in order to correctly treat recombination lines of H.

@@

Whereas SDL05 used \py itself to apply the macro-atom 
technique to a Hydrogen-only model, we
adopt a hybrid ionization and excitation scheme to model the spectra of CVs. 
Atomic species can be treated as either full macro-atoms
or using the two-level atom approximation.  This is  important in our multi-dimensional code to balance the need to calculate certain transitions in detail and the desire to minimize the set of
computational resources required.  For CVs, this allows us 
%In the latter case, they still follow the indivisible packet constraint.This allows one 
to preserve the fast treatment of the UV resonance lines while providing 
an improved representation of processes with more complex
transition paths such as recombination in H and He. 
%This is important when modeling features  such as the Balmer lines and also recombination {\sl continuum} emission.

@@

\py allows for several types of biconical wind models and various options for calculating the spectra of the components of the system. Here we describe some of the features of the  biconical wind model used in this study of CVs.

@@

Photons originate from the accretion disk and a central WD in our model. It is also possible to include
a boundary layer with user-defined temperature and emitting area. Emission from the wind is dealt with
by reprocessing of the indivisible energy packets originally produced in the disk or central WD. 
{\bf ksl -Isn't the wind also allow to emit?  If so, I think saying the wind is dealt with by reporcessing is misleading.
jm - no- all energy packets are produced by the disk and central source, and thermal emission from 
the wind would be produced by a r-packet to k-packet to r-packet path, possibly via activated macro-atoms.}

@@

\py has has some flexibility when treating the accretion disk as a source of photons. 
The disk is broken down into annuli 
such that each annulus contributes an equal amount to the bolometric luminosity. 
We adopt the temperature profile of a standard \cite{shakurasunyaev1973} $\alpha$-disk.
%shown in figure 1.
An annulus can then
be treated either as a blackbody with the corresponding effective temperature or as a stellar atmosphere model
with the appropriate surface gravity and effective temperature. 
Both of these methods have limitations. It is known
from studies of CV accretion disks that they exhibit absorption cores 
formed in an optically thick disk atmosphere.
This can impact the overall spectrum even if a wind is present 
\citep[see e.g.][]{dhillon1996}.
However, the stellar atmosphere models
have had difficulties reproducing the required spectral shape \citep{wade1988}, 
and there are inconsistencies when using a 
stellar atmosphere beneath a dense
wind region that itself goes some way to simulating the disk atmosphere. 
In addition, \cite{linnell2010} found that deviations from the predicted 
$T\propto r^{-3/4}$ temperature profile are required to fit the data
in certain cases.  {\bf ksl - I thought Wade said that both black bodies and stellar atmospheres had trouble reproducing the color of disks.  Please check. jm- this is correct. it is still relevant but needs to be worded better,
and this section needs to link with the earlier statements about e.g. Shaviv.}

@@

\begin{table}
\centering
\begin{tabular}{p{3cm}p{4cm}}
Benchmark Model \\
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{WD}$ 	 &	 $0.8 M_{\odot}$ \\ 
$R_{WD}$ 	 &	 $7\times10^{8}$cm\\ 
$\dot{M}_{acc}$ 	 &	 $10^{-8}~M_{\odot}yr^{-1}$\\ 
$\dot{M}_{wind}$  &	$10^{-9}~M_{\odot}yr^{-1}$å\\ 
$r_{min}$ 	&	 $4 R_{WD}$\\ 
$r_{max}$ 	&	 $12 R_{WD}$ \\ 
$\theta_{min}$ 	&	 $20.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $65.0^{\circ}$ \\ 
$\gamma$ 	&	 $1$ \\ 
$v_{\infty}$ 	&	 $3v_{esc}$ \\ 
%%$R_v$ 	        &	 $7\times10^{10}$cm \\ 
$R_v$ 	        &	 $100 R_{WD}$ \\ 
$\alpha$ 	&	 $1.5$ \\
\end{tabular}
\centering
\caption{
Parameters used for the geometry and kinematics of the benchmark CV model.}
\label{wind_param}
\end{table}

@@

\begin{figure} %fullpage
\includegraphics[width=\textwidth]{figures/fig5_uv_opt.eps}
\caption{
UV (left) and optical (right) synthetic spectra for model A, our benchmark model,
computed at sightlines of 10, 27.5, 45, 62.5 and 80 degrees.	
The inset plots show zoomed-in line profiles for 
\heiiuv\ and \ha. Double-peaked line emission can be seen in 
\heiiuv, \heiiopt, \ha\ and some He I lines, but the 
line emission is not always sufficient to overcome the absorption
lines from the stellar atmosphere models. The model
also produces a prominent \heiioptnew\ line at high inclinations,
}
\label{spec}
\end{figure} %fullpage\

Figure~\ref{spec} shows optical synthetic spectra for the benchmark 
model for five inclination angles ranging from $10^\circ$ to $80^\circ$ .
%5 different sightlines.  
The spectra have some lines of H and He which have emission components that are strong enough to overcome
the absorption lines in the underlying disk spectra. These emission lines are generally double peaked, 
as they are formed near the base of the wind where rotational velocity
can still dominate over the slowly increasing poloidal velocity. The line 
opacity is not yet high enough for the poloidal velocity gradient to 
cause single-peaked profiles.  {\bf ksl - So this last statement is based on information that has not yet been submitted as evidence. Be careful of the flow of the paper here.}
We also see a general trend from absorption to emission in spectral lines
with increasing inclination, as one would expect from our wind geometry.
This trend is observed in CVs and can be seen clearly in figure 1. 
% Note, however,
% that in our models the \heiiuv\ line does not exhibit this behaviour, which
% is somewhat surprising. {\bf NOTE: we need to understand this!}.

@@

The benchmark model presented above shows that a wind model
 that successfully reproduces the observed UV spectral features
of CVs,  
would also leave a clear imprint on the optical spectrum. 
It is important
to understand how changing the kinematics of the wind impacts on this
effect.   
% {\bf - ksl - I don't actually like the previous sentence.  
% We should some how say we want to study a few parameters that we think with help,
% without saying that something is beyond the scope}  
In particular, this
will help shed light on the predictions of, e.g., Murray \& Chiang (1996)
and Knigge et al. (1998) regarding the single-peaked H$\alpha$ emission
and absence of a pronounced Balmer absorption edge respectively.

@@

% A full kinematic parameter search may prove instructive, but
% is beyond the scope of this particular study. {\bf - ksl - you said this once already, and I don't think it should be explicitly stated at all}
We have conducted a limited parameter 
search, where we focus on 
two variables in order to explore the trends in our model. 
These are:

@@

\begin{table}
\centering
\begin{tabular}{p{3cm}p{4cm}}
Improved Optical Model \\
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{WD}$ 	 &	 $0.8 M_{\odot}$ \\ 
$\dot{M}_{acc}$ 	 &	 $10^{-9}~M_{\odot}yr^{-1}$\\ 
$\dot{M}_{wind}$  &	$10^{-8}~M_{\odot}yr^{-1}$\\ 
$r_{min}$ 	&	 $4 R_{WD}$\\ 
$r_{max}$ 	&	 $12 R_{WD}$ \\ 
$\theta_{min}$ 	&	 $20.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $65.0^{\circ}$ \\ 
$\gamma$ 	&	 $1$ \\ 
$v_{\infty}$ 	&	 $3v_{esc}$ \\ 
$R_v$ 	        &	 $142.9 R_{WD}$ \\ 
$\alpha$ 	&	 $4$ \\
\end{tabular}
\centering
\caption{Wind geometry parameters used in the next-generation CV model.}
\label{modelb}
\end{table}

With modest changes to the kinematics of the benchmark CV model,
it is possible to produce a model
that comes closer to reproducing the observed optical spectrum
of a NLs. The parameters for this model
are show in table~\ref{modelb}, and the synthetic spectrum
is shown in figure~\ref{uvoptb}. An in and out of eclipse comparison 
between the synthetic spectrum
at an angle of $80^\circ$ and the high-inclination NL RW Tri 
is also shown in figure~\ref{rwtricomp}. The inclination
for RW Tri has been disputed, with claimed values of $75^\circ$ \citep{groot2004},
$70.5^\circ$ \citep{smak1995}, $80^\circ$ \citep{longmore1981} and 
$82^\circ$\citep{frankking1981}. A lower inclination
would mean that our model would simply require lower densities 
and smaller emission regions to produce the required continuum level
and single peaked emission, so our conclusions from this comparison are
not particular sensitive to the value chosen.
We also emphasize that this is no sense a fit to the data. The model
is purely designed to reproduce the general behaviour of NLs
and give an indication of the likely densities and kinematics 
required in the wind.
%quotes a figure of , whereas  halevinhenden2008

The similarity with the observed spectrum is striking. 
The improved optical model produces strong emission in all the Balmer lines, 
with line to continuum ratios
comparable to those in RW Tri. 
The line to continuum ratio increases during eclipse,
as one expects if the emission is produced in a disk wind, 
thereby mimicking the behaviour of SW Sex stars.
The single-peaked \ha\ line is produced by our models with a 
very similar width to the observed spectrum, although it is perhaps slightly more
`flat-topped' in the synthesized data.

@@

\begin{figure} %fullpage
\includegraphics[width=0.8\textwidth]{figures/fig13_eclipse.eps}
\caption{{\sl Top Panel:} In and out of eclipse spectra of the high
inclination NL RW Tri. {\sl Bottom Panel:} In and out of eclipse synthetic
spectra from the improved optical model.
The artificial `absorption' feature just redward of the Balmer jump
is caused for the reasons described in section 5.2.}
\label{rwtricomp}
\end{figure} %fullpage
